Using Genetic Algorithms to Optimize the Analogue Method for Precipitation Prediction in the Swiss Alps

### 2016-12-22 Optimization case study - JoH/elsarticle-template.tex
JoH paper under correction.

The analogue method (AM) is a downscaling technique based on the idea expressed by \citet{Lorenz1956, Lorenz1969}, that similar situations in terms of atmospheric circulation are likely to lead to similar local weather \citep{Duband1970}. It uses predictor variables describing the synoptic atmospheric circulation in order to predict local-scale predictands of interest. It is often used to predict daily precipitation, either in an operational forecasting context \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Bliefernicht2010, Marty2012, Horton2012, Hamill2015, BenDaoud2016}, or a climate context \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015, Raynaud2016b}. Other predictands are also considered, such as precipitation radar images \citep{Panziera2011,Foresti2015a}, temperature \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016, Raynaud2016b}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, and radiation \citep{Bois1981, Raynaud2016b}.

@@

The method can be designed with multiple successive subsampling steps, each of them relying on different meteorological variables. A certain number of parameters define the relationship between predictors and predictands, such as the choice of the predictor variable, its pressure level and temporal window to consider, the spatial domain to use for the comparison, as well as the analogy criterion itself, and finally the number of analogue situations to keep at each subsampling level. These parameters are usually calibrated by means of a semi-automatic sequential procedure \citep[see][for the details]{Bontron2004, Radanovics2013, Horton2016a}, i.e. by optimizing each single parameter one at a time, in an arbitrarily chosen order, with no or little reassessment. This sequential approach has therefore strong limitations: (i) it cannot automatically choose the optimal pressure levels and the temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between the parameters within a level of analogy, and even less between them, and (iii) it could not easily handle new degrees of freedom, such as a weighting between the pressure levels. Thus, even if the processing involved is relatively fast, the sequential approach requires laborious assessments of predictors combinations (variables, pressure levels, temporal windows), and presents a high risk of ending in a local optimum due to subjective initial choices and lack of consideration of parameter inter-dependencies. Other calibration methods exist for specific applications, such as the radar images \citep{Panziera2011, Foresti2015a}.

@@

The analysis of the parameters cross-compatibility (Figure \ref{fig:crossing_4Zo-2MIo}) is also very similar to the one on the circulation analogy only (results for 4Zo-4MIo not shown as very similar). The same pattern can be observed, with a drop of performance for the subregions submitted to different meteorological influences. However, the losses of performance are globally more important than before, suggesting that more complex methods with moisture variables are less transposable to another subregion, even though both are located within the same grid cell of the reanalysis dataset. Moisture fields have greater variability than pressure fields, and thus a change in the spatial windows can have a greater impact on the method performance. Indeed, the two regions with the lowest cross-compatibility are the Upper Rhone Valley (7) and the Southeast ridges (8), which have similar optimal pressure levels and temporal windows to other regions, but have rather different spatial windows on the moisture predictor.

>>>>

The analogue method (AM) is a downscaling technique based on the idea expressed by \citet{Lorenz1956, Lorenz1969}, that similar situations in terms of atmospheric circulation are likely to lead to similar local weather \citep{Duband1970}. It uses predictor variables describing the synoptic atmospheric circulation in order to predict local-scale predictands of interest. It is often used to predict daily precipitation, either in an operational forecasting context \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Bliefernicht2010, Marty2012, Horton2012, Hamill2015, BenDaoud2016, Horton2016a}, or a climate context \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015, Raynaud2016b}. Other predictands are also considered, such as precipitation radar images \citep{Panziera2011,Foresti2015a}, temperature \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016, Raynaud2016b}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, and radiation \citep{Bois1981, Raynaud2016b}.

@@

The method can be designed with multiple successive subsampling steps, each of them relying on different meteorological variables. A certain number of parameters define the relationship between predictors and predictands, such as the choice of the predictor variable, its pressure level and temporal window to consider, the spatial domain to use for the comparison, as well as the analogy criterion itself, and finally the number of analogue situations to keep at each subsampling level. These parameters are usually calibrated by means of a semi-automatic sequential procedure \citep[see][for the details]{Bontron2004, Radanovics2013}, i.e. by optimizing each single parameter one at a time, in an arbitrarily chosen order, with no or little reassessment. This sequential approach has therefore strong limitations: (i) it cannot automatically choose the optimal pressure levels and the temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between the parameters within a level of analogy, and even less between them, and (iii) it could not easily handle new degrees of freedom, such as a weighting between the pressure levels. Thus, even if the processing involved is relatively fast, the sequential approach requires laborious assessments of predictors combinations (variables, pressure levels, temporal windows), and presents a high risk of ending in a local optimum due to subjective initial choices and lack of consideration of parameter inter-dependencies. Other calibration methods exist for specific applications, such as the radar images \citep{Panziera2011, Foresti2015a}.

@@

The analysis of the parameters cross-compatibility (Figure \ref{fig:crossing_4Zo-2MIo}) is also very similar to the one on the circulation analogy only (results for 4Zo-4MIo not shown as very similar). The same pattern can be observed, with a drop of performance for the subregions submitted to different meteorological influences. However, the losses of performance are globally more important than before, suggesting that more complex methods with moisture variables are less transposable to another subregion (consistent with the observations of \citet{Chardon2014}), even though both are located within the same grid cell of the reanalysis dataset. Moisture fields have greater variability than pressure fields, and thus a change in the spatial windows can have a greater impact on the method performance. Indeed, the two regions with the lowest cross-compatibility are the Upper Rhone Valley (7) and the Southeast ridges (8), which have similar optimal pressure levels and temporal windows to other regions, but have rather different spatial windows on the moisture predictor.

Predictors based on moisture variables do significantly increase the prediction skill, and are thus recommended, as long as they are reliable. In real-time forecasting, the reliability depends on the lead time: for lead times superior to 3-4 days the uncertainties related to moisture variables from NWP models are fairly high, which reduce the relevance of methods relying on this information. In climate studies, it mainly depends on the coherence of the climatologies between the archive and the model outputs. One should however not establish an AM with moisture variables for a too large region, as the transferability is reduced \citep[see also][]{Chardon2014}.
